STatistical Computing (MATH6173) Coursework 2

### 2022-12-21 draft_report.Rmd
Adding in draft of explanation of Q2

>>>>

# Question 2

## Part a
For the continuous-time random walk conditioned on $U_{0}=u_{0}$
\begin{equation}
U_{i} \sim \text{Normal}(\mu=U_{i-1},\sigma^{2}=t_{i}-t_{i-1})
\end{equation}

First by transforming $a$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(a)=Pr(X>a|X_{0}=x_{0})  \equiv Pr(U>b|U_{0}=u_{0}=p_{tail}(b))$, thus we can work with the sequence $U_{i}$ to answer equation (6). Now from equation (3):
\begin{equation*}
Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i}))=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  
\end{equation*}
For Z the standard normal and $u_{i-1}$ a sample from $U_{i-1}$
Therefore, to not waste any samples let $q_{i}$ the proposal the ith observation in the sequence $U_{i}$ be an exponential distribution shifted by $s_{i}$ where:

\begin{equation*}
 s_{i}=\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} 
\end{equation*}

As the samples $w=\{w_{1},...,w_{d}\}$ drawn from these proposals $q=\{q_{1},....,q_{d}\}$ will satisfy $w_{i}>b_{i},\hspace{1em} i \in \{1,...,d\}$ thus all values can be used to calculate $p_{tail}(b)=p_{tail}(a)$.

To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$). Thus, each iteration will need to generate these $u_{i}$ values using equation (3) the initial value $u_{0}$ and times $t_{i}i \in \{0,...,d\}$ to generate a realization $u$ of the random walk sequence $U$ and storing these in a vector to access to generate a sample from the proposal distribution.

## Part b

The algorithm will start by applying the logit transform to a and then using these transformed values b to work with values in sequence

First by transforming $\boldsymbol{a}$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b}))$, thus we can work with the sequence $U_{i}$ to answer equation (6). Therefore start the algorithm by transforming inputs into proportions (if given as percentages) and calculating $\boldsymbol{b}$.

Next we must simulate the random walk $U_{i}$ using equation 3 and storing these realizations $u_{i}$ in a vector.
Using these values and the proposals $q_{i}$ in part a generate $v_{i}$ a value for each step in sequence $U_{i}$
Calculate weights $w_{i}$ for each $v_{i}$ by formula $w_{i}=\frac{\pi_{i}(v_{i})}{Q_{i}(v_{i})})$ where $\pi_{i}$ is the pdf of $U_{i}$ and $Q_{i}$ is the density of the proposal distribution $q_{i}$
Then to calculate equation (6) for this iteration find the product of the weights as (put in equation with integral and how sampled pop is that integral)
Repeat this process until have desired n values and then calculate importance sampling estimator by finding mean of these values.
